---
title: "hanqin_visualize_data"
author: "Hanqin Du"
date: "2019/10/13"
output: html_document
editor_options:
  chunk_output_type: console
---


```{r set project folder, eval=FALSE, echo=FALSE}
par(mfrow=c(1,1))
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
getwd()
```


## Load library
```{r load library, results='hide', message=FALSE}
library(data.table)
library(tidyverse)
library(lmodel2)
library(splitstackshape)
library(gridExtra)
#library(glmnet)

```


## Load function
```{r}

mse <- function(model){
  mean(summary(model)$residuals^2)
}

plot_mean_deviation <- function(dataset,label,mean,deviation){
  plot(1:dim(dataset)[1],dataset[[mean]], pch=19,xlab="",ylab="coefficient",xaxt="n",xlim = c(0.5,dim(dataset)[1]+0.5),
       ylim=c(min(dataset[mean]-dataset[deviation]),max((dataset[mean]+dataset[deviation]))))
  lines(rbind(1:dim(dataset)[1],1:dim(dataset)[1],NA),rbind(dataset[[mean]]-dataset[[deviation]],dataset[[mean]]+dataset[[deviation]],NA))
  axis(side=1,at=1:dim(dataset)[1],labels=dataset[[label]])
}


```


## Load Data
```{r Load Data}

#ref datasets for UTRs 
UTR_raw <- read_rds("../data/Sun_mutation_UTRs.rds")
  #Get sequences from UTR_raw in a separate vector
  UTR_3 <- UTR_raw$UTR3_seq

#Load Manually created motifs list into a vector
motifs_raw <- scan("../data/list_motifs.txt", character())
motifs_cheng = c("TGTAAATA", "TGCAT", "TTTTTTA", "ATATTC")


# load Gasch's data
expressionLevel_Gasch <- read_tsv("../data/Gasch2000_complete_dataset_rename.txt", 
                       locale = locale(decimal = ","))

# convert all the expression data to number
for (i in names(expressionLevel_Gasch)[3:175]) {
  expressionLevel_Gasch[[i]] <- as.numeric(as.character(expressionLevel_Gasch[[i]]))
}

names_environment_conditions = names(expressionLevel_Gasch)[4:175]

```


## Construct Motif Frequencies Matrix from 3'UTR Ref sequences
```{r motif search}

#Dictionary for non-specific codes and converting U -> T
motifs <- motifs_raw %>% str_replace_all(c("U" = "T", "W" = "(A|T)", "S" = "(C|G)", "M" = "(A|C)", "K" = "(G|T)", "R" = "(A|G)", "Y" = "(C|T)", "B" = "(C|G|T)", "D" = "(A|G|T)", "H" = "(A|C|T)", "V" = "(A|C|G)", "N" = "(A|C|G|T)"))

#Initate ref tibble and store gene names
ref_motifs <- tibble(geneName = UTR_raw$genename)


#Search and add frequency of each c(motif) as a column in ref dataset
for (i in 1:length(motifs)){
ref_motifs <- mutate(.data = ref_motifs, 
                      !!motifs_raw[i] := str_count(UTR_3, motifs[i]))
}

names_motifs_all <- names(ref_motifs)[2:length(ref_motifs)]
```


## merge motif frequency matrix with Gasch's data
```{r merge motifs frequency with gasch data}

# merge motifs
fullTable_Gasch <- merge(expressionLevel_Gasch,ref_motifs,by = "geneName")

# remove useless motifs
motifs_count_sum <- colSums(fullTable_Gasch[names_motifs_all],na.rm=TRUE)

remove_list = NULL

for (i in names(motifs_count_sum)){
  if (motifs_count_sum[[i]] == 0){
    remove_list <- c(remove_list,i)
  }
}

names_motifs_valid <- names_motifs_all[!names_motifs_all %in% remove_list]

# convert type of motifs frequency to factor for violin plotting
fullTable_Gasch_factor = fullTable_Gasch
for(i in names_motifs_all){
    fullTable_Gasch_factor[[i]] <- as.factor(as.character(fullTable_Gasch_factor[[i]]))
}

```


## investigate the frequency sum of each motifs
```{r investigate frequency}
mean(motifs_count_sum)
median(motifs_count_sum)

plot(motifs_count_sum)
hist(motifs_count_sum, xlim = c(0,20), breaks = c(0,1,2,3,5,10,20,50,3000))

length(motifs_count_sum[motifs_count_sum == 0])
length(motifs_count_sum[motifs_count_sum < 6 & motifs_count_sum > 0])
length(motifs_count_sum[motifs_count_sum < 21 & motifs_count_sum > 5])
length(motifs_count_sum[motifs_count_sum > 20])
```


## compute correlation coefficient
```{r obtain corrilation factor}

# pick a group of temperature condition
names_temperature_condition = vector(mode="character", length=5)
names_temperature_condition[1] = "hs_17to37_20min"
names_temperature_condition[2] = "hs_21to37_20min"
names_temperature_condition[3] = "hs_25to37_20min"
names_temperature_condition[4] = "hs_29to37_20min"
names_temperature_condition[5] = "hs_33to37_20min"

# compute correlation coefficient
motif_overview <- as.data.frame(
  cor(fullTable_Gasch[names_motifs_valid],fullTable_Gasch[names_temperature_condition], use = "na.or.complete"))

# compute mean
motif_overview$R_mean <- rowMeans(motif_overview[names_temperature_condition[1:5]])

head(motif_overview)

```


## plot coefficient against each motif
```{r echo=FALSE, fig.width=18, fig.height=18}

# plot multi-line to check if any motif sensitive to temperature

ggplot(data = melt(setDT(as.data.frame(motif_overview[1:50,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  labs(x = "ESR", y = "correlation coefficient")

p1 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[1:5,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) + 
  labs(x = "ESR", y = "correlation coefficient")

p2 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[6:10,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) + 
  labs(x = "ESR",y = "correlation coefficient")

p3 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[11:15,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) + 
  labs(x = "ESR",y = "correlation coefficient")

p4 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[16:20,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p5 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[21:25,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p6 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[26:30,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p7 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[31:35,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p8 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[36:40,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p9 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[41:45,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

p10 <- ggplot(data = melt(setDT(as.data.frame(motif_overview[46:50,names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + 
  geom_line(aes(colour=rn)) + 
  ylim(-0.25,0.1) +
  labs(x = "ESR",y = "correlation coefficient")

grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10, ncol=2)


## plot motifs with high correlation coefficient
ggplot(data = melt(setDT(as.data.frame(motif_overview[c('UGUAHMNUA','TGTATAWT','ATATTC'),names_temperature_condition]), keep.rownames = TRUE), "rn"),
       aes(x=variable,y=value,group = rn)) + geom_line(aes(colour=rn)) + ylim(-0.25,0.1)

```


## plot the distribution of the motifs with high coefficient 
```{r warning=FALSE, echo=FALSE, fig.width=10}

# apply filter |x| > 0.05 to check significant value
#correlation_filter <- motif_overview[names_temperature_condition]
#correlation_filter[correlation_filter<0.05&correlation_filter>-0.05] <- NA


# plot violin graph to check distribution
grid.arrange(
  ggplot(fullTable_Gasch_factor,aes(x=UGUAHMNUA,y=hs_17to37_20min)) + geom_violin(),
  ggplot(fullTable_Gasch_factor,aes(x=TGTATAWT,y=hs_17to37_20min)) + geom_violin(),
  ggplot(fullTable_Gasch_factor,aes(x=ATATTC,y=hs_17to37_20min)) + geom_violin(),
  ncol=3
  )
```


## carry out linear regression on two set of expression data
```{r}

# train models
formula_lm_17to37 <- as.formula(paste("hs_17to37_20min ~",paste(names_motifs_valid, collapse = "+")))
lm_17to37 = lm(formula_lm_17to37,fullTable_Gasch)

formula_lm_25to37 <- as.formula(paste("hs_25to37_20min ~",paste(names_motifs_valid, collapse = "+")))
lm_25to37 = lm(formula_lm_25to37,fullTable_Gasch)


# organize linear model's data and merge them into motif_overview 
coefficients_lm_17to37 = as.data.frame(summary(lm_17to37)$coefficients)
colnames(coefficients_lm_17to37)[colnames(coefficients_lm_17to37)=="Estimate"] <- "W_17to37"
colnames(coefficients_lm_17to37)[colnames(coefficients_lm_17to37)=="Std. Error"] <- "SE_17to37"

coefficients_lm_25to37 = as.data.frame(summary(lm_25to37)$coefficients)
colnames(coefficients_lm_25to37)[colnames(coefficients_lm_25to37)=="Estimate"] <- "W_25to37"
colnames(coefficients_lm_25to37)[colnames(coefficients_lm_25to37)=="Std. Error"] <- "SE_25to37"


motif_overview <- merge(motif_overview,
              cbind(coefficients_lm_17to37[c('W_17to37','SE_17to37')],coefficients_lm_25to37[c('W_25to37','SE_25to37')]),
              by = "row.names")

motif_overview <- transform(motif_overview, row.names=Row.names)
colnames(motif_overview)[colnames(motif_overview)=="Row.names"] <- "motif"

head(motif_overview)

```


## plot each coefficient with their deviation
```{r warning=FALSE, fig.width=15, fig.height=15}

# coefficient from linear model: "hs_17to37_20min"
par(mfrow=c(2,2))
plot(lm_17to37)
mse(lm_17to37)

par(mfrow=c(5,1))
{plot_mean_deviation(motif_overview[1:10,],"motif","W_17to37","SE_17to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[11:20,],"motif","W_17to37","SE_17to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[21:30,],"motif","W_17to37","SE_17to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[31:40,],"motif","W_17to37","SE_17to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[41:50,],"motif","W_17to37","SE_17to37")
  abline(h=0, col="red")}


# coefficient from linear model: "hs_25to37_20min"
par(mfrow=c(2,2))
plot(lm_25to37)
mse(lm_25to37)

par(mfrow=c(5,1))
{plot_mean_deviation(motif_overview[1:10,],"motif","W_25to37","SE_25to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[11:20,],"motif","W_25to37","SE_25to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[21:30,],"motif","W_25to37","SE_25to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[31:40,],"motif","W_25to37","SE_25to37")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[41:50,],"motif","W_25to37","SE_25to37")
  abline(h=0, col="red")}

```


## plot distribution of the motifs with unexpected high coefficient
```{r echo=FALSE, warning=FALSE}
# motif with unexpected high coefficient
ggplot(fullTable_Gasch_factor,aes(x=WAAAGGTAGTAAGT,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=TTTTCTAGGDD,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=WWTMGTATATTGTMA,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=RAGGAGACGT,y=hs_17to37_20min)) + geom_violin() + geom_point()

```

## plot distribution of the motifs expected and actual have a high coefficient
```{r echo=FALSE, warning=FALSE}
# motif expected and have a high coefficient
ggplot(fullTable_Gasch_factor,aes(x=UGUAHMNUA,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=ATATTC,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=TGTATAWT,y=hs_17to37_20min)) + geom_violin() + geom_point()


```

## plot distribution of the motifs expected and actual have a high coefficient
```{r echo=FALSE, warning=FALSE}
# motif expected and have a low coefficient
ggplot(fullTable_Gasch_factor,aes(x=HWNCAUUWY,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=AGACAATCT,y=hs_17to37_20min)) + geom_violin() + geom_point()

```


## carry out linear regression with high-frequency motif (>20)
```{r}

# filter out motif with frequency lower than 20
remove_list = NULL

for (i in names(motifs_count_sum)){
  if (motifs_count_sum[[i]] < 20){
    remove_list <- c(remove_list,i)
  }
}

names_motifs_valid <- names_motifs_all[!names_motifs_all %in% remove_list]


# train models
formula_lm_17to37 <- as.formula(paste("hs_17to37_20min ~",paste(names_motifs_valid, collapse = "+")))
lm_17to37 = lm(formula_lm_17to37,fullTable_Gasch)

formula_lm_25to37 <- as.formula(paste("hs_25to37_20min ~",paste(names_motifs_valid, collapse = "+")))
lm_25to37 = lm(formula_lm_25to37,fullTable_Gasch)


# organize linear model's data and merge them into motif_overview 
coefficients_lm_17to37 = as.data.frame(summary(lm_17to37)$coefficients)
colnames(coefficients_lm_17to37)[colnames(coefficients_lm_17to37)=="Estimate"] <- "W_17to37_filtered"
colnames(coefficients_lm_17to37)[colnames(coefficients_lm_17to37)=="Std. Error"] <- "SE_17to37_filtered"

coefficients_lm_25to37 = as.data.frame(summary(lm_25to37)$coefficients)
colnames(coefficients_lm_25to37)[colnames(coefficients_lm_25to37)=="Estimate"] <- "W_25to37_filtered"
colnames(coefficients_lm_25to37)[colnames(coefficients_lm_25to37)=="Std. Error"] <- "SE_25to37_filtered"


motif_overview <- merge(motif_overview,
              cbind(coefficients_lm_17to37[c('W_17to37_filtered','SE_17to37_filtered')],
                    coefficients_lm_25to37[c('W_25to37_filtered','SE_25to37_filtered')]),
              all = TRUE,by = "row.names")

motif_overview <- motif_overview[-1,-1]
motif_overview <- transform(motif_overview, row.names=motif)

head(motif_overview)

```

## plot each coefficient with their deviation
```{r fig.width=15, fig.height=15}
#plot(fullTable_Gasch$WUUGUAWUWU,fullTable_Gasch$`heat shock 17 to 37, 20 minutes`)
#plot(fullTable_Gasch$AAAATAAAG,fullTable_Gasch$`heat shock 17 to 37, 20 minutes`)

par(mfrow=c(2,2))
plot(lm_17to37)
mse(lm_17to37)

par(mfrow=c(2,1))
{plot_mean_deviation(motif_overview[names_motifs_valid[1:10],],"motif","W_17to37_filtered","SE_17to37_filtered")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[names_motifs_valid[11:21],],"motif","W_17to37_filtered","SE_17to37_filtered")
  abline(h=0, col="red")}


par(mfrow=c(2,2))
plot(lm_25to37)
mse(lm_25to37)

par(mfrow=c(2,1))
{plot_mean_deviation(motif_overview[names_motifs_valid[1:10],],"motif","W_25to37_filtered","SE_25to37_filtered")
  abline(h=0, col="red")}
{plot_mean_deviation(motif_overview[names_motifs_valid[11:21],],"motif","W_25to37_filtered","SE_25to37_filtered")
  abline(h=0, col="red")}


```

## plot distribution of the motifs with unexpected high coefficient
```{r echo=FALSE, warning=FALSE}
# apply filter |x| > 0.05 to check significant value
motif_overview_filter <- motif_overview
motif_overview_filter[motif_overview_filter<0.05&motif_overview_filter>-0.05] <- NA

# motif with unexpected high coefficient

ggplot(fullTable_Gasch_factor,aes(x=WUUGUAWUWU,y=hs_17to37_20min)) + geom_violin() + geom_point()
ggplot(fullTable_Gasch_factor,aes(x=TGTAAATA,y=hs_17to37_20min)) + geom_violin() + geom_point()

```





